Coupling PDEs on XD-YD domains with Lagrange multipliers

### 2019-05-21 stab/stab.tex
the discrete coupled surface formulation

@@

  \draw[black, very thick] (2, 0) -- (2, 4);
  \draw[->, black] (2, 1) -- (3, 1);
  \node[fill=white, opacity=0.9, above] at (2.5, 1) {$n$};
  \node[fill=white, opacity=0.9, right] at (2.025, 2) {$\Gamma$};
\end{tikzpicture}
    \end{center}
    %\vspace{-20pt}
    \caption{Schematic domain of coupled multiscale problem.}
    \label{fig:coupled_domain}
    %\vspace{5pt}    
\end{figure}
\end{minipage}
%
Let $\Omega=\Omega^{+}\cup\Omega^{-}$ and $\Gamma_D=\Gamma^{+}_D\cup\Gamma^{-}_N$.
Introducing a Lagrange multiplier $p=\jump{\nabla u \cdot n}\in Q=(H^{-1/2}_{00}(\Gamma))^{\prime}$
the weak form of \eqref{eq:coupled} reads: Find $u\in V=H^1_{0, \Gamma_D}(\Omega)$,
$\hat{v}\in\hat{V}=H^1_{0, \partial\Gamma}(\Gamma)$, $p\in Q$ such that
%
\begin{equation}\label{eq:coupled_weak}
  \begin{aligned}
    &\int_{\Omega}\nabla u\cdot \nabla v &\phantom{+\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &+\int{\Gamma}p v &= \int_{\Omega}f v \quad &\forall v\in V,\\
    &\phantom{\int_{\Omega}\nabla u\cdot \nabla v} &+{\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &-\int{\Gamma}p\hat{v} &= \int_{\Gamma}\hat{f}\hat{v} \quad &\forall \hat{v}\in \hat{V},\\
    &\int_{\Gamma}u q &-\int_{\Gamma}\hat{u}q &\phantom{-\int{\Gamma}q\hat{v}} &= \int_{\Gamma}h q \quad &\forall q\in Q.
  \end{aligned}
\end{equation}
%
\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[scale=0.60]
      % Minus
  \draw[black, very thick] (1,0) rectangle (4, 3);
  \draw[black, very thick] (1,0) -- (4, 3);
  \node[fill=white, opacity=0.9] at (2.5, 1.5) {$\Omega^{-}_{H}$};

@@

>>>>

\newcommand{\Miro}[1]{\textcolor{cyan}{Miro: #1}}

@@

  \draw[black, very thick] (2, 0) -- (2, 4);
  \draw[->, black] (2, 1) -- (3, 1);
  \node[fill=white, opacity=0.9, above] at (2.5, 1) {$n$};
  \node[fill=white, opacity=0.9, right] at (2.025, 2) {$\Gamma$};
\end{tikzpicture}
    \end{center}
    %\vspace{-20pt}
    \caption{Schematic domain of coupled multiscale problem.}
    \label{fig:coupled_domain}
    %\vspace{5pt}    
\end{figure}
\end{minipage}
%
Let $\Omega=\Omega^{+}\cup\Omega^{-}$ and $\Gamma_D=\Gamma^{+}_D\cup\Gamma^{-}_N$.
Introducing a Lagrange multiplier $p=\jump{\nabla u \cdot n}\in Q=(H^{-1/2}_{00}(\Gamma))^{\prime}$
the weak form of \eqref{eq:coupled} reads: Find $u\in V=H^1_{0, \Gamma_D}(\Omega)$,
$\hat{v}\in\hat{V}=H^1_{0, \partial\Gamma}(\Gamma)$, $p\in Q$ such that
%
\begin{equation}\label{eq:coupled_weak}
  \begin{aligned}
    &\int_{\Omega}\nabla u\cdot \nabla v &\phantom{+\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &+\int_{\Gamma} p v &= \int_{\Omega}f v \quad &\forall v\in V,\\
    &\phantom{\int_{\Omega}\nabla u\cdot \nabla v} &+{\int_{\Gamma}\nabla \hat{u}\cdot \nabla \hat{v}} &-\int_{\Gamma} p\hat{v} &= \int_{\Gamma}\hat{f}\hat{v} \quad &\forall \hat{v}\in \hat{V},\\
    &\int_{\Gamma}u q &-\int_{\Gamma}\hat{u}q &\phantom{-\int{\Gamma}q\hat{v}} &= \int_{\Gamma}h q \quad &\forall q\in Q.
  \end{aligned}
\end{equation}
%
\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[scale=0.60]
      % Minus
  \draw[black, very thick] (1,0) rectangle (4, 3);
  \draw[black, very thick] (1,0) -- (4, 3);
  \node[fill=white, opacity=0.9] at (2.5, 1.5) {$\Omega^{-}_{H}$};

@@

\section{Coupled multiscale problem in 3 mesh formulation}\label{sec:coupled_3}
The discrete formulations thus far concerned only two meshes, $\Omega_H$ and $\Gamma_h$,
and are in this sense similar to Problem 2 of the 3$d$-1$d$ problem. Here we
shall consider a 2$d$ ``analogue'' of Problem 1. In particular the reduced
domain and the coupling/Lagrange multiplier domain shall be different.

Let us recall that in Problem 1 (of the coupled 3$d$-1$d$ problem) the coupling surface $\Gamma$ isn't necessarily
identical/related to the cylindrical surface whose centerline is the reduced
domain $\Gamma$. We propose to take advantage of the fact that (to certain extent?)
the surface may be chosen freely in the following construction which automatically
satisfies the requirement of conformity of the triangulation of $\Omega$ to the
multiplier surface.

Let $\Omega_H$, $\Gamma_h$ be the two \emph{independent} discretizations of the
bulk and the centerline. In particular, $\Gamma_h$ is not assumed to consist only
of the edges of the tetrahedra in $\Omega_H$. However, we assume that the $\Gamma_h$
is finer than $\Omega_H$ restricted to the line. Next, let $\mathcal{K}_H$ be
the set of all tetrahedra in $\Omega_H$ intersected by elements of $\Gamma_h$.
Finally, the coupling surface $\omega_H$ shall consist of those facets
of elements in $\mathcal{K}_H$ which are \emph{not} intersected by the curve.
The process in the 2$d$ setting of Figure \ref{fig:coupled_domain} is shown
in Figure \ref{fig:mesh3}. We remark that the coupling surface changes with
discretization.

%
\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[scale=0.60]
      % Minus
  \draw[black, very thick] (0,0) rectangle (3, 3);
  \draw[black, very thick] (0,0) -- (3, 3);
  \node[left] at (0, 1.5) {$\Omega^{-}_{H}$};

  \draw[black, very thick] (3,0) -- (6, 0);
  \draw[black, very thick] (3,0) -- (6, 3);
  \draw[black, very thick] (3,3) -- (6, 3);  
  % Right
    \draw[black, very thick] (6,0) rectangle (9, 3);
  \draw[black, very thick] (6,0) -- (9, 3);
  \node[right] at (9, 1.5) {$\Omega^{+}_{H}$};

  \node[mark size=3pt, color=black] at (6, 0) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=black] at (9, 0) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=black] at (6, 3) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=black] at (9, 3) {\pgfuseplotmark{*}};

  \draw[blue, very thick] (3,0) -- (3, 3);
  \draw[blue, very thick] (6,0) -- (6, 3);
  \node[mark size=3pt, color=blue] at (6, 0) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=blue] at (6, 3) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=blue] at (3, 0) {\pgfuseplotmark{*}};
  \node[mark size=3pt, color=blue] at (3, 3) {\pgfuseplotmark{*}};

  \node[below] at (3, 0) {$\omega_{H}$};
  \node[below] at (6, 0) {$\omega_{H}$};

  \end{tikzpicture}
    \end{center}
  \caption{Coupling surface $\omega_H$ consists of facets of elements
  of $\Omega_H$ not intersected by $\Gamma_h$.}
    \label{fig:mesh3}
\end{figure}

To investigate the idea we consider the multiscale problem \eqref{eq:coupled_weak}.
Note that its left hand side defines an operator
%
\[
\begin{pmatrix}
  A & & T^{\prime}\\
  & \hat{A} & -I\\
  T & -I & 0\\
\end{pmatrix},
\]
where $A$, $\hat{A}$ are the Laplacians on $\Omega$ and $\Gamma$ respectively
and $T$, $I$ are the trace and identity operators. Let now $V_H=H^{1}_0(\Omega)$,
$\hat{V}_h=H^{1}_0(\Gamma_h)$, $Q=(H^{-1/2}_{00}(\omega_H))^{\prime}$. We shall
consider a modified problem defined by operator
%
\begin{equation}\label{eq:mesh3}
\begin{pmatrix}
  A & & T^{\prime}\\
  & \hat{A} & -E^{\prime}_{\omega}\\
  T_{\omega} & -E_{\omega} & 0\\
\end{pmatrix},
\end{equation}
%
with $T_{\omega}: V\rightarrow Q^{\prime}$ (trace) and $E_{\omega}: \hat{V}\rightarrow Q^{\prime}$
(extension)
\[
\langle T_{\Omega} u, q\rangle = \int_{\omega_H} uq, \quad\quad
%
\langle E_{\Omega} \hat{u}, q\rangle = \int_{\omega_H} \hat{u} q.
\]
%

In the following experiments we investigate stability and approximation
properties of \eqref{eq:mesh3}. To this end we consider a Riesz map preconditioner
based on spaces $V$, $\hat{V}$ and $Q$ and the manufactured test cased from \S \ref{sec:coupled}.
The spaces are all discretized with P1 elements and we have $H=3h$. As before, we enforce boundary
conditions on $Q$. The extension operator is defined as extension by constant
from the nearest degree of freedom of $\hat{V}$.

\begin{table}
  \begin{center}
    \scriptsize{
  \begin{tabular}{c|ccc|c|c}
    \hline
    $h$ & $\norm{u-u_H}_1$ & $\norm{\hat{u}-\hat{u}_h}_1$ & $\norm{p-p_h}_0$ & $n$ & $\kappa$\\
    \hline
8.58E-2 & 1.49E0(--)    & 6.74E-1(--)   & 7.22E0(--)   & 31 & 13   \\
4.35E-2 & 9.89E-1(0.60) & 3.46E-1(0.98) & 7.59E0(-0.07)& 40 & 22 \\
2.19E-2 & 6.66E-1(0.58) & 1.74E-1(1.00) & 7.37E0(0.04) & 48 & 41 \\
1.10E-2 & 4.57E-1(0.55) & 8.75E-2(1.00) & 7.25E0(0.02) & 58 & 79 \\
5.51E-3 & 3.17E-1(0.53) & 4.37E-2(1.00) & 7.14E0(0.02) & 75 & 154 \\
2.76E-3 & 2.22E-1(0.52) & 2.18E-2(1.00) & 7.09E0(0.01) & 75 & 304 \\
1.38E-3 & 1.56E-1(0.51) & 1.09E-2(1.00) & 7.06E0(0.01) & 77 & -- \\
    \hline
  \end{tabular}
    }
    \caption{Three mesh formulation. Approximation errors, number of preconditioned MinRes iterations ($n$) and
      condition number of the preconditioned problem \eqref{eq:coupled_weak} ($\kappa$) discretized
      with P1-P1-P1 elements.}
  \label{tab:mesh3}
  \end{center}
\end{table}
%

Table \ref{tab:mesh3} summarizes the preliminary results. With the exception of the multiplier
convergence in all the variables can be seen. However, we recall that the
true solution is here defined as $\jump{\nabla u\cdot n}$ on $\Gamma$ while
the discrete one lives on $\omega_H$ so there is at the very least some
geometrical error. Observing the condition numbers we also see that the
formulation is unstable.

\subsection{TODO}
\begin{itemize}
\item Build \eqref{eq:mesh3} for just half of the domain, so $\omega_H$ is connected
\item How is this problem related 2$d$-2$d$-2$d$ problem with ``thin'' (middle) domain
  which becomes $\Gamma$?
\item What are then the approximation properties?
\item Stabilization
\end{itemize}
